A fast algorithm for finding efficient decisions in multi-objective problems with black-box multiextremal functions

### 2022-01-13 2022/OM&S/paper.tex
abstract

\begin{abstract}
This template is for authors who are preparing a manuscript for a Taylor \& Francis journal using the \LaTeX\ document preparation system and the \texttt{interact} class file, which is available via selected journals' home pages on the Taylor \& Francis website.
\end{abstract}

@@

In general case, there exist three key factors determining the complexity of analyzing these models. The crucial feature corresponds to the typical case where criteria are contradictory: increasing one criterion leads to decreasing the other. Such inconsistent behavior generates the necessity of introducing a compromise in the joint analysis of criteria and consideration of a set of compromised (partial) solutions as a complete solution of a multiobjective problem (Pareto set), see, e.g., \cite{Collette2004,Ehrgott2005}.
The second complexity factor is associated with the dimensionality or number of varied parameters of the problem because increasing the problem dimension leads to significant growth of computational cost.

@@

In the next section we describe an information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.

@@

The Lipschitzian functions are, in general case, multiextremal. If we consider the convolution (\ref{conv}) aggregating multiextremal criteria it will be multiextremal as well.

In the computational scheme proposed in the paper we consider a family $S$ of scalar problems (\ref{conv_problem}) that are solved jointly and take into account the information obtained in the course of already completed optimizations. For this purpose in the region of weight coefficients determined by the conditions (\ref{lambda}) a uniform grid is built and the family $S$ consists of the problems (\ref{conv_problem}) with coefficients $\lambda$ corresponding the nodes of the grid.

@@

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ which are scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}). This procedure will be called Repeated Algorithm with Local Refinements and Data Accumulation (RALR-DA).

@@

During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
\textcolor[rgb]{1,0.41,0.13}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.

@@

\begin{figure}
\centering
\includegraphics[width=0.6\textwidth]{fig1.png} 
\caption{Contour plots of 2-dimensional convolution.}\label{fig1} 
\end{figure}

The Pareto front in one of test problems is shown in Fig.~\ref{fig2}.

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig2.png} 
\caption{Contour plots of 2-dimensional convolution.}\label{fig2} 
\end{figure}

The results of the experiment are presented in Tab.~\ref{tab1} and Fig.~\ref{fig3} where Pareto corresponds to the estimation obtained by the grid method.

@@

As for metaheuristic algorithms, their results differ in different experiments.  For example, in 4-dimensinal bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria. On the whole, as the most stable metaheuristic method, the algorithm OMOPSO can be recognized.

>>>>

\begin{abstract}
The paper is devoted to consideration of multiextremal black-box MCO problems with time-consuming criteria and proposes a new computational algorithm for solving such problems. This algorithm is based on combination of convolution approach, dimensionality reduction, and information-statistical global optimization algorithm. For approximating the Pareto set a family of maximum convolutions is considered generating problems of scalar multiextremal optimization. The latters are reduced by means of Peano mappings to the equivalent problems of univariate optimization that are solved by a modification of the information-statistical global optimization method. This method applies the technique of local refinements and uses in the course of current univariate optimization the information about criteria that was obtained when solving completed optimization problems. These features allow one to accelerate the search of global solution.  
The efficiency of the proposed MCO algorithm is experimentally estimated on representative classes of multidimensional multiextremal test problems. Comparison with 5 nature-inspired particle swarm and genetic algorithms on the base of the hypervolume index reflecting quality of Pareto set evaluation demonstrates qualitative results of the proposed method.
\end{abstract}

@@

In general case, there exist three key factors that determine the complexity of analyzing these models. The crucial feature corresponds to the typical case where criteria are contradictory: increasing one criterion leads to decreasing the other. Such inconsistent behavior generates the necessity of introducing a compromise in the joint analysis of criteria and consideration of a set of compromised (partial) solutions as a complete solution of a multiobjective problem (Pareto set), see, e.g., \cite{Collette2004,Ehrgott2005}.
The second complexity factor is associated with the dimensionality or number of varied parameters of the problem because increasing the problem dimension leads to significant growth of computational cost.

@@

In the next section propose a modified information-statistical algorithm with accelerated convergence and its application to searching for Pareto optimal solutions in multiextremal MCO problems.

@@

The Lipschitzian functions are, in general case, multiextremal. If we consider the convolution (\ref{conv}) aggregating multiextremal criteria it is multiextremal as well.

In the computational scheme proposed in the paper we consider a family $S$ of scalar problems (\ref{conv_problem}) that are solved jointly and take into account the information obtained in the course of already completed optimizations. For this purpose in the region of weight coefficients determined by the conditions (\ref{lambda}) a uniform grid is built and the family $S$ consists of the problems (\ref{conv_problem}) with coefficients $\lambda$ corresponding to the grid nodes.

@@

Finalizing the proposed general scheme  let us describe a procedure  of successive solving the problems of the family $S$ which are scalar problems (\ref{conv_problem}) for different weight vectors $\lambda = (\lambda_1, \dots,\lambda_k)$ (\ref{lambda}). This procedure will be called Reduced Algorithm with Local Refinements and Data Accumulation (RALR-DA).

@@

During all experiments the method ALR was used with parameters $r=5.6$, $\sigma = 18$, $T=3$. 
\textcolor[rgb]{1,0,0}{The use of $T=3$ corresponds to alternation of one ``local'' iteration and two ``global'' iterations.}
Accuracy $\varepsilon$ from the stopping condition (\ref{eq:23}) was $0.05$ in bi-criteria problems and $0.01$ in 4-criteria ones.

@@

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig1.png} 
\caption{Contour plots of 2-dimensional convolution.}\label{fig1} 
\end{figure}

\textcolor[rgb]{1,0,0}{The Pareto front in one of test problems is shown in Fig.~\ref{fig2a} and Fig.~\ref{fig2b}. Here Fig.~\ref{fig2a} corresponds the Pareto front computed on the uniform grid with the step 0.05 in the search domain whereas Fig.~\ref{fig2b} shows the Pareto front computed by RALR-DA. In the last case it is required only 350 trials to obtain the solution with the same accuracy.}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig2a.png} 
\caption{Pareto front computed by grid technique.}\label{fig2a} 
\end{figure}

\begin{figure}
\centering
\includegraphics[width=0.7\textwidth]{fig2b.png} 
\caption{Pareto front computed by RALR-DA.}\label{fig2b} 
\end{figure}

\textcolor[rgb]{1,0,0}{The averaged HV indices obtained while solving the two-dimensional bi-criteria problems are presented} in Tab.~\ref{tab1} and Fig.~\ref{fig3} where ``Pareto'' corresponds to the estimation obtained by the grid method.

@@

As for metaheuristic algorithms, their results differ in different experiments.  For example, in 4-dimensional bi-criteria case the method SPEA is the best among nature-inspired algorithms, but it is the worst in 2-dimensional problems with 2 criteria. On the whole, as the most stable metaheuristic method, the algorithm OMOPSO can be recognized.
